Can Auxiliary Indicators Improve COVID-19 Forecasting and Hotspot Prediction?

### 2021-10-29 forecast/paper/paper.tex
small changes to intro/discussion

% Please add a significance statement to explain the relevance of your work
\significancestatement{Validated forecasting methodology should be a key 
  element in the public health response to any fast-moving epidemic or 
  pandemic. A widely-used model for predicting the future spread of a temporal
  process, based on past data, is an autoregressive (AR) model. While basic,
  such an AR model (properly trained) can already be competitive with top models
  in operational use for COVID-19 forecasting. In this paper, we 
  exhibit five auxiliary indicators---based on de-identified medical insurance
  claims, self-reported symptoms via online surveys, and COVID-related Google
  searches---that further improve the predictive accuracy of an AR model in
  COVID-19 forecasting. The most substantial gains appear to be in quiescent  
  times; but the Google search indicator appears to also offer improvements
  during upswings in pandemic activity.   
}

@@

In the companion paper \cite{Reinhart:2021}, we analyze correlations between
various indicators and COVID case rates. These correlations are natural
summaries of the contemporaneous association between an indicator and COVID
cases, but they fall short of delivering a satisfactory answer to the question
that motivates the current article: is the information contained in an indicator
demonstrably useful for the prediction tasks we care about? For such a question,
lagged correlations (e.g., measuring the correlation between an indicator and
COVID case rates several days in the future) move us in the right direction, but
still do not move us all the way there. The question about \textit{utility for
  prediction} is focused on a much higher standard than simply asking about 
correlations; to be useful in forecast or hotspot models, an indicator must
provide relevant information that is not otherwise contained in past values of
the case rate series itself. We assess this by inspecting the difference in
predictive performance of simple autoregressive models trained with and without
access to past values of a particular indicator.

@@

\subsection{Effects of Leading or Lagging Behavior}

@@

Can auxiliary indicators improve COVID-19 forecasting and hotspot prediction
models?  Our answer, based on analyzing five auxiliary indicators from the
COVIDcast API (defined using from medical insurance claims, internet-based
surveys, and internet search trends) is undoubtedly ``yes''. However, there are
levels of nuance to such an answer that must be explained.  None of the
indicators that we have investigated revolutionize the tractability of the prediction
problem, rendering it easy when it was once hard (in the absence of auxiliary
information).  Rather, the gains in accuracy from the indicator models (over an
autoregressive model based only on past case rates) appear to be nontrivial, and
consistent across modes of analysis, but modest.  In forecasting, the indicator
models are found to be most useful in periods in which case rates are flat or
trending down, rather than periods in which case rates are trending up (as one
might hope to see is the benefit provided by a hypothetical ``leading
indicator'').

@@

In this paper, we have chosen to consider only very simple forecasting models,
while devoting most of our effort to accounting for as much of the complexity of
the underlying data and evaluation as possible.  In fact, our paper is as much
about demonstrating how one might address questions about model comparisons and
evaluation in forecasting and hotspot prediction in general, as it is about
providing rigorous answers to such questions in the context of COVID-19 case
rates in particular.  We hope that others will leverage our framework, and build
on it, so that it can be used to guide work that advances the frontier of
predictive modeling for epidemics and pandemics.

>>>>

% Please add a significance statement to explain the relevance of your work
\significancestatement{Validated forecasting methodology should be a key 
  element in the public health response to any fast-moving epidemic or 
  pandemic. A widely-used model for predicting the future spread of a temporal
  process, based on past data, is an autoregressive (AR) model. While basic,
  such an AR model (properly trained) is already competitive with top models
  in operational use for COVID-19 forecasting. In this paper, we 
  exhibit five auxiliary indicators---based on de-identified medical insurance
  claims, self-reported symptoms via online surveys, and COVID-related Google
  searches---that further improve the predictive accuracy of an AR model in
  COVID-19 forecasting. The most substantial gains appear to be in quiescent  
  times; but the Google search indicator appears to also offer improvements
  during upswings in pandemic activity.   
}

@@

In the companion paper \cite{Reinhart:2021}, we analyze correlations between
various indicators and COVID case rates. These correlations are natural
summaries of the contemporaneous association between an indicator and COVID
cases, but they fall short of delivering a satisfactory answer to the question
that motivates the current article: is the information contained in an indicator
demonstrably useful for the prediction tasks we care about?
% For such a question,
% lagged correlations (e.g., measuring the correlation between an indicator and
% COVID case rates several days in the future) move us in the right direction, but
% still do not move us all the way there.
Demonstrating \textit{utility for
  prediction} is a much higher standard than simply asking about 
correlations; to be useful in forecast or hotspot models, an indicator must
provide relevant information that is not otherwise contained in past values of
the case rate series itself (see the pioneering work on Granger Causality
\cite{engle1987a,Granger:1969} as well as detailed discussion below). We
assess this directly by inspecting the difference in  
predictive performance of simple autoregressive models trained with and without
access to past values of a particular indicator.

@@

\subsection{Effects of Leading or Lagging Behavior}
\label{sec:leading-lagging}

@@

Can auxiliary indicators improve COVID-19 forecasting and hotspot prediction
models?  Our answer, based on analyzing five auxiliary indicators from the
COVIDcast API (defined using from medical insurance claims, internet-based
surveys, and internet search trends) is undoubtedly ``yes''. However, there are
levels of nuance to such an answer that must be explained.  None of the
indicators that we have investigated are transformative, rendering the
prediction problem  easy when it was once hard (in the absence of auxiliary
information).
% This seems like a place to reference Superforecasters (or related) if we're
% going to do it. ``The expectation should never be
% that there is a silver bullet. Too much solid work refutes that. Again lack of
% prior art citations to some very solid prior empirical work, and repetition of
% a long discredited Hedgehog/Silver bullet strawman.''  
Rather, the gains in accuracy from the indicator models (over an
autoregressive model based only on past case rates) appear to be nontrivial, and
consistent across modes of analysis, but modest.  In forecasting, the indicator
models are found to be most useful in periods in which case rates are flat or
trending down, rather than periods in which case rates are trending up (as one
might hope to see is the benefit provided by a hypothetical ``leading
indicator'').

@@

In this paper, we have chosen to consider only very simple forecasting models,
while devoting most of our effort to accounting for as much of the complexity of
the underlying data and evaluation as possible.
% The below was held out as particularly anathema to the PNAS ethos.
%
% In fact, our paper is as much
% about demonstrating how one might address questions about model comparisons and
% evaluation in forecasting and hotspot prediction in general, as it is about
% providing rigorous answers to such questions in the context of COVID-19 case
% rates in particular.  We hope that others will leverage our framework, and build
% on it, so that it can be used to guide work that advances the frontier of
% predictive modeling for epidemics and pandemics.
%
% Added by DJM
Despite their simplicity, these models are competitive with top forecasting
models submitted to the COVID-19 Forecast Hub in real time, and they are
demonstrably robust to the careful scrutiny undertaken here.
